Cradles of the first stars: self-shielding, halo masses, and multiplicity

### 2019-11-15 paper.tex
Adding referee_2.tex to begin to respond. I have adjusted the paper according to some of their first comments. I have indicated which ones in the .tex file.

If within \textbf{10} pc, multiple cells meet this criteria, then a single Pop III star forms at the center of mass of these cells. \textbf{In a parameter study, we have found that the number of Pop III star particles is insensitive to this algorithmic merging distance.} The mass of the Pop III star is randomly sampled from an initial mass function (IMF) of the form:
\begin{equation} \label{IMF}
	f(\log M)dM = M^{-1.3} \exp \left[-\left( \frac{M_{\rm char}}{M}\right)^{1.6} \right]dM
\end{equation}
where M$_{\rm char}$ = 20 M$_{\odot}$ \citep{Hirano17}. This characteristic mass is different from the W12 choice of 100 M$_{\odot}$. This results in an exponential cutoff below M$_{\rm char}$ and a power-law IMF above. \textbf{The Pop III IMF mass ranges from $1 \leq M / M_{\odot} \leq 300$. }Once the cell meets the above criteria and the stellar mass is chosen, an equal amount of gas is removed from the grid in a sphere which contains twice the stellar mass that is centered on the new star particle. The new Pop III star then gains the mass-weighted velocity of the gas contained within that sphere. In this simulation, we do not track any small scale fragmentation which may form low mass stars \citep{Greif11_P3Cluster}. We are only tracking massive Pop III star formation.

@@

In redshift bins of $\Delta z$ = 1, the mean halo mass hosting Pop III stars is determined, and plotted as black dots against the redshift bins in Figure \ref{fig:mean_mass}. The median for each bin is represented by a cross, and the 15.9 and 84.1 percentiles are plotted below and above the mean, respectively. The M01 relation (Equation \ref{mthresh}) for our LW background intensity in Eq. \ref{LWbg} is shown as the black solid line. Notably, the mean halo mass falls well below the relation, at $<10^{6}$ M$_{\odot}$ until $z = 15$ (270 Myr), at which point, the mean halo mass steadily rises above $3 \times 10^{6}$ M$_{\odot}$. The large discrepancy between our data and the mass threshold from M01 is indicative of the \hh{} shielding which is included in our simulation. \hh{} shielding allows for halos to cool at much lower masses, and therefore, Pop III stars are forming in these halos at earlier times. In the M01 analysis, \hh{} shielding is neglected in their calculations for a variety of reasons, including the Doppler broadening of LW bands and large column densities of \hh{} only beginning to form at late times. Interestingly, a full radiation hydrodynamics simulation, such as ours, does not produce identical results as M01 predicts. Throughout the literature, there appears to be a consensus that \hh{} self-shielding will help smaller mass halos collapse in the presence of a LW background radiation \citep[e.g.][]{Yoshida03, Ricotti01, Glover01, Hartwig15}. As can be seen by the red solid line in Figure \ref{fig:mean_mass}, nearly 100\% of the halos hosting Pop III stars fall below the M01 relation, until around a redshift of 15, at which point, the mean halo mass begins to rise above the relation. The increase in the mean halo mass above the M01 relation is indicative of the increased LW intensity permeating the box due to a galaxy that becomes very active in star formation at late times. This galaxy has a halo mass of $3.6 \times 10^{8} \rm M_{\odot}$, a total stellar mass of $3.1 \times 10^{6} \rm M_{\odot}$, and has a peak Pop III SFR of $2.7 \times 10^{-4}\, \rm M_{\odot}\ yr^{-1}$ at $z=15$ (250 Myr) and a peak metal enriched SFR of $5.4 \times 10^{-2}\, \rm M_{\odot} \ yr^{-1}$ at $z=9.5$ (500~Myr). This galaxy produces a large amount of LW radiation which dominates the box, and drives up the minimum halo mass required for the formation of Pop III stars. This increase in the LW background can also be seen in Figure \ref{fig:JLW_xe_mass}, where by the end of the simulation, the LW background rises to greater than 1 $J_{21}$.

@@

There has been a large amount of work done throughout the literature investigating the relationship between the LW background intensity and halo masses. \citet{Tegmark97} numerically integrated a chemical network to calculate the minimum mass a halo must have in order to cool, by considering \hh{} formation and cooling rates. They conclude that cooling by \hh{} is efficient and leads to the first formation of structures. They found a minimum halo mass needed to collapse which depends on the virial temperature of the halo and the virial redshift. At $z = 15$, this minimum halo mass is $10^{6} M_{\odot}$. \citet{Trenti09} uses a similar argument as \citet{Tegmark97}, and found that metal-free halos can exist until z $\approx$ 6 using cosmological simulations and an analytical model for metal enrichment. They also provide some insight into whether or not Pop III supernova rates are high enough within these metal-free halos to allow for the LSST to see these regions of the universe. They used the minimum halo mass capable of cooling via \hh{} from \citet{Trenti09_SFR} as one part of their halo mass model (see the blue solid line in Figure \ref{fig:compare_JLW_mass}, given our LW background). Neither \citet{Trenti09} nor \citet{Tegmark97} included the effects of \hh{} self-shielding, which can explain the discrepancy between our data and their mass threshold relationship (Figure \ref{fig:compare_JLW_mass}). For a LW background of $J_{\rm LW} / J_{21} = 0.2$, we see halos hosting Pop III stars down to $4 \times 10^{5}$ M$_{\odot}$, while the relation from \citet{Trenti09_SFR} would expect to see minimum halo masses capable of cooling, and thus capable of hosting Pop III stars, at $10^7$ M$_{\odot}$, almost two orders of magnitude larger. Since we include \hh{} self-shielding, Pop III stars are allowed to form in smaller mass halos for a given LW background. \textbf{\citet{Visbal14} used one-zone models to study the gaseous centers of dark matter halos and how they react to a LW background. They use a particular parametrization of the central gas density, which then follows their analytic equation for critical mass (see their Equation 4) to within a factor of two. This critical mass is plotted in our Figure \ref{fig:compare_JLW_mass} at redshift 20. Here almost all of our halos fall below their critical mass relationship, with only the highest mass halos rising above. \citeauthor{Visbal14} also compare their results to three-dimensional hydrodynamical simulations, where they find similar threshold halo masses as \citet{Machacek01}.} \citet{Mebane18} used a semi-analytic model of star formation, including feedback properties such as a LW background, photoionization due to Pop III stars, supernovae of Pop III stars and metal-enriched stars, and chemical enrichment, to determine for how long Pop III stars will survive for and found that Pop III stars can continue to form until z $\approx$ 6. They found a minimum halo mass for hosting Pop III star formation, but with the caveat that they did not include \hh{} self-shielding. They found that when Pop III stars contribute to the LW background, the minimum halo mass for Pop III star formation is $4 \times 10^{6} M_{\odot}$ at $z = 20$, similar to results from \citet{Tegmark97}. The results from \citet{Trenti09} and \citet{Mebane18} have higher halo masses in comparison with M01, although this more closely matches results from simulations \citep[see][]{Wise07_UVB, OShea08}. It should be noted that because our box is fairly small, we cannot capture the cosmic variance of rare halos and galaxies. For example, at late times, a single galaxy dominates the LW radiation field, and drives up the host halo masses. However, we would expect this to happen at different times in other cosmological volumes. Therefore, we cannot directly compare the time dependence, however, a comparison as a function of LW intensity is still valid.

\citet{Yoshida03} used cosmological simulations to study the formation of primordial star-forming clouds. They followed the growth of structure to find where gas cools and condenses which would form the first stars, and how the effects of LW radiation may affect these gas clouds. Importantly, they included \hh{} self-shielding within halos. In a series of simulations, they found a minimum halo mass for those halos hosting gas clouds which may result in active star formation (see their Figure 12). They found that in the presence of a LW background of 0.01 J$_{21}$, the minimum halo mass is $7 \times 10^{5} M_{\odot}$. When \hh{} self-shielding is taken into account along with a LW background, their minimum halo mass decreases to $4.5 \times 10^{5} M_{\odot}$. This value lies close to the case where there is no LW background applied, where the minimum halo mass is $3.5 \times 10^{5} M_{\odot}$. They found that \hh{} self-shielding does appear to be an efficient mechanism enabling primordial gas cooling. In comparison with our work, we find similar minimum halo masses for each redshift, although we do see a wider range of halo masses, ranging from $2.8 \times 10^{5}$ M$_{\odot}$ to $1.4 \times 10^{7}$ M$_{\odot}$. While our halos never experience a LW intensity as low as 0.01 J$_{21}$, we find that their minimum mass for this LW intensity lies in the same mass range as our halos (see blue open circle in Figure \ref{fig:compare_JLW_mass}). \citet{Wise07_UVB} used cosmological simulations to investigate \hh{} cooling in a LW background. They found that \hh{} cooling is dominant even when there is a large LW background present. They did not include self-shielding and subsequently found halo masses at their collapse that lie well above the M01 relation (see red open triangles in Figure \ref{fig:compare_JLW_mass}). Their J$_{\rm LW}$ = 0 control is plotted in Figure \ref{fig:compare_JLW_mass} at J$_{\rm LW}$/J$_{21}$ = 10$^{-4}$. \citet{OShea08} used cosmological simulations to investigate Pop III star formation in various LW backgrounds. They found that due to an increased LW background, there is a delay in star formation, and thus there is an increase in the halo masses at collapse. They also ignored \hh{} self-shielding in their calculations which can again account for the increased halo masses they found compared to this work, as can be seen in Figure \ref{fig:compare_JLW_mass} (open green squares). Their halo masses are similar to the results of \citet{Wise07_UVB}. Their control of J$_{21}$ = 0 control is plotted at J$_{LW}$/J$_{21}$ = 10$^{-4}$. \textbf{\citet{Hirano15} used cosmological simulations to study 1540 collapsing metal-free gas clouds in the early universe to derive the corresponding primordial stellar mass distribution. They find that stars form in halos above virial masses $M_{\rm vir} = 2.1 \times 10^5 M_{\odot}$ at $z = 30$ and $M_{\rm vir} = 9.9 \times 10^5 M_{\odot}$ at $z = 10$. These masses are consistent with our results with our mean halo mass lying between these values.}

@@

We also do not include streaming velocities between baryons and dark matter within our simulation \citep{Tselia11, Greif11_Delay, Naoz12, OLeary12}. Streaming velocities can suppress star formation at very high ($z \ga 20$) redshifts only, because the streaming velocities decrease with time. Pop III star formation is delayed by an average of $\delta z = 4$ by increasing the halo mass needed to overcome the bulk velocity by about a factor of three \citep{Greif11_Delay}. The streaming velocities will generally result in lower gas densities within halos as well as an offset of the peak density from the center of the halo \citep{OLeary12}. Often in the literature, streaming velocities are only studied for a few halos. But \citet{Naoz13} estimated the minimum mass needed to retain the bulk of the baryons within the dark matter halo using a series of cosmological simulations, the results of which can be seen in Figure \ref{fig:compare_JLW_mass} (see shaded bands). These halos only reach up to about $2 - 3 \times 10^5 M_\odot$ when the streaming velocity is about $1\sigma_{\rm vbc}$, which is smaller than the star forming halos in our simulation. \citet{Naoz13} mention that at the largest streaming velocity, their simulations do not match expected values due to a small sample size of high mass halos. Our results indicate that halos collapse at larger masses that are not as affected by typical streaming velocities ($<1\sigma_{\rm vbc}$).  In the evolution of such a halo, it would have started to accumulate gas above these critical masses while also being irradiated by a LW background.  Afterwards, the collapse will be controlled by the processes discussed in Section \ref{sec:variations}, and thus streaming velocities would likely not affect our results. \textbf{\citet{Schauer19} give minimum and average halo masses necessary for enough cold dense gas to form in halo centers to support star formation for various streaming velocities in their study of star formation in minihalos and its relation to the EDGES 21-cm absorption feature \citep{EDGES}. Ranging from 0 $\rm v_{\rm bc}$ to 3.0 $\rm v_{\rm bc}$ in units of $\sigma_{\rm rms}$, they found that the minimum halo mass ranges from $5 \times 10^5 M_{\odot}$ to $10^7 M_{\odot}$, and the average halo mass ranges from about $2 \times 10^6 M_{\odot}$ to $4 \times 10^7 M_{\odot}$. Our halo masses generally fall within these ranges, confirming again that only the rarest and strongest of the streaming velocities would affect our results.}

>>>>

If within \textbf{10} pc, multiple cells meet this criteria, then a single Pop III star forms at the center of mass of these cells. \textbf{In a parameter study, we have found that the number of Pop III star particles is insensitive to this algorithmic merging distance. We found that even when this number was decreased to 0.065 pc, the number of stars which formed was unchanged.} The mass of the Pop III star is randomly sampled from an initial mass function (IMF) of the form:
\begin{equation} \label{IMF}
	f(\log M)dM = M^{-1.3} \exp \left[-\left( \frac{M_{\rm char}}{M}\right)^{1.6} \right]dM
\end{equation}
where M$_{\rm char}$ = 20 M$_{\odot}$ \citep{Hirano17}. This characteristic mass is different from the W12 choice of 100 M$_{\odot}$. This results in an exponential cutoff below M$_{\rm char}$ and a power-law IMF above. \textbf{The Pop III IMF mass ranges from $1 \leq M / M_{\odot} \leq 300$. }Once the cell meets the above criteria and the stellar mass is chosen, an equal amount of gas is removed from the grid in a sphere which contains twice the stellar mass that is centered on the new star particle. The new Pop III star then gains the mass-weighted velocity of the gas contained within that sphere. In this simulation, we do not track any small scale fragmentation which may form low mass stars \citep{Greif11_P3Cluster}. We are only tracking massive Pop III star formation.

@@

In redshift bins of $\Delta z$ = 1, the mean halo mass hosting Pop III stars is determined, and plotted as black dots against the redshift bins in Figure \ref{fig:mean_mass}. The median for each bin is represented by a cross, and the 15.9 and 84.1 percentiles are plotted below and above the mean, respectively. The M01 relation (Equation \ref{mthresh}) for our LW background intensity in Eq. \ref{LWbg} is shown as the black solid line. Notably, the mean halo mass falls well below the relation, at $<10^{6}$ M$_{\odot}$ until $z = 15$ (270 Myr), at which point, the mean halo mass steadily rises above $3 \times 10^{6}$ M$_{\odot}$. The large discrepancy between our data and the mass threshold from M01 is indicative of the \hh{} shielding which is included in our simulation. \hh{} shielding allows for halos to cool at much lower masses, and therefore, Pop III stars are forming in these halos at earlier times. In the M01 analysis, \hh{} shielding is neglected in their calculations for a variety of reasons, including the Doppler broadening of LW bands and large column densities of \hh{} only beginning to form at late times. Interestingly, a full radiation hydrodynamics simulation, such as ours, does not produce identical results as M01 predicts.

Throughout the literature, there appears to be a consensus that \hh{} self-shielding will help smaller mass halos collapse in the presence of a LW background radiation \citep[e.g.][]{Yoshida03, Ricotti01, Glover01, Hartwig15}. As can be seen by the red solid line in Figure \ref{fig:mean_mass}, nearly 100\% of the halos hosting Pop III stars fall below the M01 relation, until around a redshift of 15, at which point, the mean halo mass begins to rise above the relation. The increase in the mean halo mass above the M01 relation is indicative of the increased LW intensity permeating the box due to a galaxy that becomes very active in star formation at late times. This galaxy has a halo mass of $3.6 \times 10^{8} \rm M_{\odot}$, a total stellar mass of $3.1 \times 10^{6} \rm M_{\odot}$, and has a peak Pop III SFR of $2.7 \times 10^{-4}\, \rm M_{\odot}\ yr^{-1}$ at $z=15$ (250 Myr) and a peak metal enriched SFR of $5.4 \times 10^{-2}\, \rm M_{\odot} \ yr^{-1}$ at $z=9.5$ (500~Myr). This galaxy produces a large amount of LW radiation which dominates the box, and drives up the minimum halo mass required for the formation of Pop III stars. This increase in the LW background can also be seen in Figure \ref{fig:JLW_xe_mass}, where by the end of the simulation, the LW background rises to greater than 1 $J_{21}$.

@@

There has been a large amount of work done throughout the literature investigating the relationship between the LW background intensity and halo masses. \citet{Tegmark97} numerically integrated a chemical network to calculate the minimum mass a halo must have in order to cool, by considering \hh{} formation and cooling rates. They conclude that cooling by \hh{} is efficient and leads to the first formation of structures. They found a minimum halo mass needed to collapse which depends on the virial temperature of the halo and the virial redshift. At $z = 15$, this minimum halo mass is $10^{6} M_{\odot}$. \citet{Trenti09} uses a similar argument as \citet{Tegmark97}, and found that metal-free halos can exist until z $\approx$ 6 using cosmological simulations and an analytical model for metal enrichment. They also provide some insight into whether or not Pop III supernova rates are high enough within these metal-free halos to allow for the LSST to see these regions of the universe. They used the minimum halo mass capable of cooling via \hh{} from \citet{Trenti09_SFR} as one part of their halo mass model (see the blue solid line in Figure \ref{fig:compare_JLW_mass}, given our LW background). Neither \citet{Trenti09} nor \citet{Tegmark97} included the effects of \hh{} self-shielding, which can explain the discrepancy between our data and their mass threshold relationship (Figure \ref{fig:compare_JLW_mass}). For a LW background of $J_{\rm LW} / J_{21} = 0.2$, we see halos hosting Pop III stars down to $4 \times 10^{5}$ M$_{\odot}$, while the relation from \citet{Trenti09_SFR} would expect to see minimum halo masses capable of cooling, and thus capable of hosting Pop III stars, at $10^7$ M$_{\odot}$, almost two orders of magnitude larger. Since we include \hh{} self-shielding, Pop III stars are allowed to form in smaller mass halos for a given LW background.

\textbf{\citet{Visbal14} used one-zone models to study the gaseous centers of dark matter halos and how they react to a LW background. They use a particular parametrization of the central gas density, which then follows their analytic equation for critical mass (see their Equation 4) to within a factor of two. This critical mass is plotted in our Figure \ref{fig:compare_JLW_mass} at redshift 20. Here almost all of our halos fall below their critical mass relationship, with only the highest mass halos rising above. \citeauthor{Visbal14} also compare their results to three-dimensional hydrodynamical simulations, where they find similar threshold halo masses as \citet{Machacek01}.} \citet{Mebane18} used a semi-analytic model of star formation, including feedback properties such as a LW background, photoionization due to Pop III stars, supernovae of Pop III stars and metal-enriched stars, and chemical enrichment, to determine for how long Pop III stars will survive for and found that Pop III stars can continue to form until z $\approx$ 6. They found a minimum halo mass for hosting Pop III star formation, but with the caveat that they did not include \hh{} self-shielding. They found that when Pop III stars contribute to the LW background, the minimum halo mass for Pop III star formation is $4 \times 10^{6} M_{\odot}$ at $z = 20$, similar to results from \citet{Tegmark97}. The results from \citet{Trenti09} and \citet{Mebane18} have higher halo masses in comparison with M01, although this more closely matches results from simulations \citep[see][]{Wise07_UVB, OShea08}.

It should be noted that because our box is fairly small, we cannot capture the cosmic variance of rare halos and galaxies. For example, at late times, a single galaxy dominates the LW radiation field, and drives up the host halo masses. However, we would expect this to happen at different times in other cosmological volumes. Therefore, we cannot directly compare the time dependence, however, a comparison as a function of LW intensity is still valid.

\citet{Yoshida03} used cosmological simulations to study the formation of primordial star-forming clouds. They followed the growth of structure to find where gas cools and condenses which would form the first stars, and how the effects of LW radiation may affect these gas clouds. Importantly, they included \hh{} self-shielding within halos. In a series of simulations, they found a minimum halo mass for those halos hosting gas clouds which may result in active star formation (see their Figure 12). They found that in the presence of a LW background of 0.01 J$_{21}$, the minimum halo mass is $7 \times 10^{5} M_{\odot}$. When \hh{} self-shielding is taken into account along with a LW background, their minimum halo mass decreases to $4.5 \times 10^{5} M_{\odot}$. This value lies close to the case where there is no LW background applied, where the minimum halo mass is $3.5 \times 10^{5} M_{\odot}$. They found that \hh{} self-shielding does appear to be an efficient mechanism enabling primordial gas cooling. In comparison with our work, we find similar minimum halo masses for each redshift, although we do see a wider range of halo masses, ranging from $2.8 \times 10^{5}$ M$_{\odot}$ to $1.4 \times 10^{7}$ M$_{\odot}$. While our halos never experience a LW intensity as low as 0.01 J$_{21}$, we find that their minimum mass for this LW intensity lies in the same mass range as our halos (see blue open circle in Figure \ref{fig:compare_JLW_mass}).

\citet{Wise07_UVB} used cosmological simulations to investigate \hh{} cooling in a LW background. They found that \hh{} cooling is dominant even when there is a large LW background present. They did not include self-shielding and subsequently found halo masses at their collapse that lie well above the M01 relation (see red open triangles in Figure \ref{fig:compare_JLW_mass}). Their J$_{\rm LW}$ = 0 control is plotted in Figure \ref{fig:compare_JLW_mass} at J$_{\rm LW}$/J$_{21}$ = 10$^{-4}$. \citet{OShea08} used cosmological simulations to investigate Pop III star formation in various LW backgrounds. They found that due to an increased LW background, there is a delay in star formation, and thus there is an increase in the halo masses at collapse. They also ignored \hh{} self-shielding in their calculations which can again account for the increased halo masses they found compared to this work, as can be seen in Figure \ref{fig:compare_JLW_mass} (open green squares). Their halo masses are similar to the results of \citet{Wise07_UVB}. Their control of J$_{21}$ = 0 control is plotted at J$_{LW}$/J$_{21}$ = 10$^{-4}$. \textbf{\citet{Hirano15} used cosmological simulations to study 1540 collapsing metal-free gas clouds in the early universe to derive the corresponding primordial stellar mass distribution. They find that stars form in halos above virial masses $M_{\rm vir} = 2.1 \times 10^5 M_{\odot}$ at $z = 30$ and $M_{\rm vir} = 9.9 \times 10^5 M_{\odot}$ at $z = 10$. These masses are consistent with our results with our mean halo mass lying between these values.}

@@

We also do not include streaming velocities between baryons and dark matter within our simulation \citep{Tselia11, Greif11_Delay, Naoz12, OLeary12}. Streaming velocities can suppress star formation at very high ($z \ga 20$) redshifts only, because the streaming velocities decrease with time. Pop III star formation is delayed by an average of $\delta z = 4$ by increasing the halo mass needed to overcome the bulk velocity by about a factor of three \citep{Greif11_Delay}. The streaming velocities will generally result in lower gas densities within halos as well as an offset of the peak density from the center of the halo \citep{OLeary12}. Often in the literature, streaming velocities are only studied for a few halos. But \citet{Naoz13} estimated the minimum mass needed to retain the bulk of the baryons within the dark matter halo using a series of cosmological simulations, the results of which can be seen in Figure \ref{fig:compare_JLW_mass} (see shaded bands). These halos only reach up to about $2 - 3 \times 10^5 M_\odot$ when the streaming velocity is about $1\sigma_{\rm vbc}$, which is smaller than the star forming halos in our simulation. \citet{Naoz13} mention that at the largest streaming velocity, their simulations do not match expected values due to a small sample size of high mass halos. Our results indicate that halos collapse at larger masses that are not as affected by typical streaming velocities ($<1\sigma_{\rm vbc}$).  In the evolution of such a halo, it would have started to accumulate gas above these critical masses while also being irradiated by a LW background.  Afterwards, the collapse will be controlled by the processes discussed in Section \ref{sec:variations}, and thus streaming velocities would likely not affect our results.

\textbf{\citet{Schauer19} give minimum and average halo masses necessary for enough cold dense gas to form in halo centers to support star formation for various streaming velocities in their study of star formation in minihalos and its relation to the EDGES 21-cm absorption feature \citep{EDGES}. Ranging from 0 $\rm v_{\rm bc}$ to 3.0 $\rm v_{\rm bc}$ in units of $\sigma_{\rm rms}$, they found that the minimum halo mass ranges from $5 \times 10^5 M_{\odot}$ to $10^7 M_{\odot}$, and the average halo mass ranges from about $2 \times 10^6 M_{\odot}$ to $4 \times 10^7 M_{\odot}$. Our halo masses generally fall within these ranges, confirming again that only the rarest and strongest of the streaming velocities would affect our results.}
